# Risk of Bias Assessment (ROBINS-I)

```{r}
#| include: false
for (p in c("R/_common.R", "../_common.R", "book/R/_common.R")) if (file.exists(p)) { source(p); break }
```

::: infobox
**Downloads:** [ROBINS-I Data (xlsx)](../data/results_rob.xlsx){download="data_ROBINS_I.xlsx"} · [ROBINS-I R-Script](../R/mod_RoB.R){download="mod_ROBINS_I.R"}
:::

```{r rob-data}
# -------------------------
# Load data
# -------------------------
data_RoB <- readxl::read_xlsx(
  path  = es_path("book/data/results_rob.xlsx"),
  sheet = "Overall",
  range = "A1:H10"
)

data_ConFac <- readxl::read_xlsx(
  path  = es_path("book/data/results_rob.xlsx"),
  sheet = "ConFac",
  range = "A1:G10"
)

data_ConFac_notes <- tribble(
  ~Item,   ~Note,
  "D1.1",  "No exercise 24h-48h pre-intervention?",
  "D1.2",  "Standardized food intake among participants (e.g. all fasted or standardized breakfast)?",
  "D1.3",  "No caffeine, acute medication, or alcohol 24h before exercise?",
  "D1.4",  "Identical warm-up between participants?",
  "D1.5",  "good way to assess the controlled intensity (HR control, VO2, VT)",
  "D1.6",  "Homogeneity of the group (sex, age, BMI, training status)"
)
```

```{r rob-plot}
#| label: rob-plot
#| fig-width: 14
#| fig-height: 10
#| out-width: 100%

# =========================================================================
# SETTINGS (editable vectors)
# =========================================================================

## 1. Study order (alphabetical descending: Z->A, A at the bottom)
study_order <- sort(unique(data_RoB$Study), decreasing = TRUE)  # Z->A

## 2. Plot C - domain order (editable, "Overall" at the bottom)
domain_order_C <- c(
  "Bias due to confounding",
  "Bias due to selection of participants",
  "Bias due to deviations from intended interventions",
  "Bias due to missing data",
  "Bias in measurement of outcomes",
  "Bias in selection of the reported result",
  "Overall risk of bias"  # <-- bottom
)

## 3. Bar order in Plot C (Low -> Moderate -> Serious, left to right)
risk_levels_bar <- c("Low", "Moderate", "Serious")
risk_levels_all <- c("Low", "Moderate", "Serious", "Critical")

## 4. Colours (robvis standard)
robvis_cols <- c(
  Low      = "#00C000",
  Moderate = "#E6E600",
  Serious  = "#E31A1C",
  Critical = "#7F0000"
)

## 5. Show 'Critical' level?
show_critical <- FALSE
active_levels <- if (show_critical) risk_levels_all else risk_levels_bar
active_cols   <- robvis_cols[active_levels]

## 6. Font sizes
base_font   <- 12
legend_font <- 10

# =========================================================================
# DOMAIN LABELS
# =========================================================================
domain_labels <- c(
  Overall = "Overall risk of bias",
  D1      = "Bias due to confounding",
  D3      = "Bias due to selection of participants",
  D4      = "Bias due to deviations from intended interventions",
  D5      = "Bias due to missing data",
  D6      = "Bias in measurement of outcomes",
  D7      = "Bias in selection of the reported result"
)

# =========================================================================
# THEMES
# =========================================================================
common_theme <- theme(
  text         = element_text(size = base_font, color = "black"),
  axis.text    = element_text(size = base_font),
  legend.text  = element_text(size = legend_font),
  legend.title = element_text(size = legend_font),
  plot.title   = element_text(face = "bold", size = base_font + 2, hjust = 0),
  plot.tag     = element_text(face = "bold", size = base_font + 4)
)

top_axis_theme <- theme(
  panel.grid        = element_blank(),
  axis.text         = element_text(color = "black"),
  axis.text.x       = element_text(margin = margin(b = 6)),
  axis.ticks        = element_blank(),
  axis.line         = element_blank(),
  axis.line.x.top   = element_line(color = "black", linewidth = 0.8),
  axis.line.y       = element_blank(),
  plot.margin       = margin(10, 10, 10, 10),
  legend.position   = "bottom",
  legend.direction  = "horizontal",
  legend.background = element_blank(),
  legend.key        = element_blank()
)

# =========================================================================
# PLOT A: RoB matrix (points per study x domain), study order Z->A
# =========================================================================
df_matrix <- data_RoB |>
  pivot_longer(cols = c(D1, D3, D4, D5, D6, D7, Overall),
               names_to = "Domain", values_to = "Risk") |>
  mutate(
    Risk   = factor(trimws(Risk), levels = risk_levels_all),
    Domain = factor(Domain, levels = c("D1", "D3", "D4", "D5", "D6", "D7", "Overall")),
    Study  = factor(Study, levels = study_order)  # Z->A
  )

p_A <- ggplot(df_matrix, aes(x = Domain, y = Study)) +
  geom_point(aes(fill = Risk), shape = 21, size = 9,
             color = "black", stroke = 1, show.legend = TRUE) +
  scale_fill_manual(
    values = robvis_cols,
    breaks = risk_levels_all,
    drop   = !show_critical
  ) +
  scale_x_discrete(position = "top") +
  coord_equal() +
  labs(x = NULL, y = NULL, fill = NULL) +
  theme_minimal() +
  top_axis_theme

# =========================================================================
# PLOT B: Confounding-factor matrix (points per study x item), synced order
# =========================================================================
df_conf_long <- data_ConFac |>
  pivot_longer(cols = -Study, names_to = "Item", values_to = "Answer") |>
  mutate(
    Item   = factor(Item, levels = data_ConFac_notes$Item),
    Study  = factor(Study, levels = study_order),  # Z->A (synced with A)
    Answer = factor(Answer, levels = c("Unsure", "Sure"))
  )

p_B <- ggplot(df_conf_long, aes(x = Item, y = Study)) +
  geom_point(aes(fill = Answer), shape = 21, size = 9,
             color = "black", stroke = 1, show.legend = TRUE) +
  scale_x_discrete(position = "top") +
  scale_fill_manual(
    values = c(Unsure = "orange1", Sure = "#00C000"),
    breaks = c("Sure", "Unsure"),
    drop   = TRUE
  ) +
  coord_equal() +
  labs(x = NULL, y = NULL, fill = NULL) +
  theme_minimal() +
  top_axis_theme

# =========================================================================
# PLOT C: Risk-of-bias proportional bar chart
# Domain order: Overall at the bottom; bar order Low -> Moderate -> Serious
# =========================================================================
risk_levels_C <- c("Low", "Moderate", "Serious")
active_cols_C <- robvis_cols[risk_levels_C]

df_long_C <- data_RoB |>
  pivot_longer(
    cols = c(D1, D3, D4, D5, D6, D7, Overall),
    names_to = "Domain",
    values_to = "Risk"
  ) |>
  mutate(
    Risk = factor(trimws(Risk), levels = risk_levels_C),
    Domain = recode(Domain, !!!domain_labels),
    # rev() needed because of coord_flip() so "Overall" ends up at the bottom
    Domain = factor(Domain, levels = rev(domain_order_C))
  ) |>
  filter(!is.na(Risk))

p_C <- ggplot(df_long_C, aes(x = Domain, fill = Risk)) +
  geom_bar(
    position = position_fill(reverse = TRUE),
    width = 0.55,
    color = "black",
    linewidth = 0.5
  ) +
  coord_flip() +
  scale_y_continuous(
    limits = c(0, 1),
    breaks = c(0, .25, .50, .75, 1),
    labels = percent_format(accuracy = 1),
    expand = c(0, 0)
  ) +
  scale_fill_manual(
    values = active_cols_C,
    breaks = risk_levels_C,
    limits = risk_levels_C,
    drop   = FALSE
  ) +
  guides(
    fill = guide_legend(
      reverse = FALSE,
      override.aes = list(colour = "black")
    )
  ) +
  labs(x = NULL, y = NULL, fill = NULL) +
  theme_classic() +
  theme(
    panel.grid      = element_blank(),
    axis.line.y     = element_blank(),
    axis.line.x     = element_line(color = "black", linewidth = 0.8),
    axis.ticks.y    = element_blank(),
    axis.ticks.x    = element_line(color = "black", linewidth = 0.8),
    legend.position = "bottom",
    legend.key      = element_rect(fill = "white", colour = "black"),
    plot.margin     = margin(10, 30, 10, 10)
  )

# =========================================================================
# FINAL LAYOUT: A + B on top, C below
# =========================================================================
final_plot <- ((p_A | p_B) + plot_layout(widths = c(7, 6)) & common_theme) /
  (p_C & common_theme) +
  plot_layout(heights = c(2, 1.2)) +
  plot_annotation(tag_levels = list(c("A", "B", "C")))

final_plot
```

::: infobox
**Confounding-factor items (Panel B).**

-   **D1.1** — No exercise 24–48 h pre-intervention?
-   **D1.2** — Standardized food intake among participants (e.g. all fasted or standardized breakfast)?
-   **D1.3** — No caffeine, acute medication, or alcohol 24 h before exercise?
-   **D1.4** — Identical warm-up between participants?
-   **D1.5** — Good way to assess the controlled intensity (HR control, VO₂, VT)?
-   **D1.6** — Homogeneity of the group (sex, age, BMI, training status)?
:::